---
title: "Step 1: Data Preparation"
author: "Ceri Ann Williams"
date: "6/3/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Required packages
```{r, include=FALSE}
library(jsonlite)
library(dplyr) #used
library(tidyr)
library(DT)
library(ggplot2) # used
#library(stringr)
#library(data.table)
library(readr) # used
```

#Load and filter data
The Yelp Datasets can be downloaded here:
When loading the data in the code below:
- "review.json" refers to the Reviews dataset
- "business.json" refers to the Business dataset
- "yelp_user" refers to the User dataset
```{r}
#load data sets
##review
yelp_reviews <- stream_in(con = file("review.json"))
reviews <- flatten(yelp_reviews)
rm(yelp_reviews)
#number of observations: 4,736,897
reviews$date <- as.Date(reviews$date)

#filter dates: 21.10.2015 - 26.01.2017
subset_date <- reviews %>% filter(date >= as.Date("2015-10-21") & date <= as.Date("2017-01-26"))
#number of observations: 1,276,961

##business
yelp_business <- stream_in(con = file("business.json"))
business <- flatten(yelp_business)
rm(yelp_business)
#number of observations: 156,639

##user
yelp_user <- read_csv(file = "yelp_user.csv")
#number of observations: 1,326,100

#merge 3 data sets
total <- merge(business, reviews, by="business_id")
#number of observations: 4,736,897
final <- merge(total, yelp_user, by="user_id")
#number of observations: 4,700,938

#filter by date
subset_date2 <- final %>% filter(date >= as.Date("2015-10-21") & date <= as.Date("2017-01-26"))
#number of observations: 1,268,595
summary(subset_date2$date)

#filter category: "restaurants"
#flatten categories vector
final$categories <- vapply(final$categories, paste, collapse = ", ", character(1L))
restaurant_reviews <- subset(subset_date2, grepl("Food|Restaurants", subset_date2$categories))
#number of observations: 845,764

#take random subset of 15,000 observations
set.seed(3107)
final <- restaurant_reviews[sample(1:nrow(restaurant_reviews), 15000, replace=FALSE), ]
```

The above taken sample will be used for further analysis
```{r}
sample <- read_csv(file="Dataset1.csv")
colnames(sample)
```

#Add DV - Helpfulness
Binary variable:
- 0 if not helpful (= no useful votes) 
- 1 if helpful (= at least one useful vote)
```{r}
#Overall
sample$DV <- ifelse(sample$useful.x == 0, 0, 1)
t1 <- table(sample$DV)
prop.table(t1)

#Distribution of reviews in this time frame
summary(sample$date2)
ggplot(sample, aes(x = date2)) +
  geom_histogram()
```

#Save subsample
```{r}
#final$categories <- vapply(final$categories, paste, collapse = ", ", character(1L))
#write.csv(final, file="subsamplefinal_1.csv")
```

#Open sample
```{r}
#mysample <- read_csv(file = "/Users/Ceri/Desktop/Thesis Data/Final Dataset/junesubsamplefinal_1.csv")
```

#Descriptives of subsample
##Distribution of DV
- Distribution of DV
  Not Helpful: 9,770 (65.1%)
  Helpful: 5,230 (34.9%)
```{r}
sample %>% group_by(DV) %>% summarise(n=n()) %>% mutate(freq = n / sum(n))
```

##Overall
- Total number of reviews: 15,000
- Total number of businesses: 9,305
- Total number of users: 13,774
- Timeframe: 21.10.2015 - 26.01.2017
```{r}
sample %>% group_by(review_id) %>% summarise(count=n())
sample %>% group_by(business_id) %>% summarise(count=n())
sample %>% group_by(user_id) %>% summarise(count=n())
max(sample$date2)
min(sample$date2)
```

##Review descriptives
-	Distribution of star ratings of reviews
```{r}
ggplot(sample, aes(x = stars.x)) +
  geom_histogram()
```

-	Average number of useful votes per review: 0.8482
```{r}
summary(sample$useful.x)
```

#Variable selection
Filter only variables needed for further analysis
```{r}
sample <- sample %>% select(review_id, user_id, DV, text, stars.y, date2, review_count.y, elite, fans, yelping_since, compliment_hot, compliment_more, compliment_profile, compliment_cute, compliment_list, compliment_note, compliment_photos, compliment_writer, compliment_funny, compliment_cool, compliment_plain)

colnames(sample)
```

##New variable: sum of all compliments a user received
- Sum of compliments a user received: reviewer_compliments
```{r}
#sample$reviewer_compliments <- sample$compliment_cool + sample$compliment_cute + sample$compliment_funny + sample$compliment_hot + sample$compliment_list + sample$compliment_more + sample$compliment_note + sample$compliment_plain + sample$compliment_photos + sample$compliment_profile + sample$compliment_writer

#colnames(sample)
```

#Save data
```{r}
write.csv(sample, file="Dataset2.csv")
```
