---
title: "Step 2: Feature Engineering"
author: "Ceri Ann Williams"
date: "6/3/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr) 
library(textcat) 
library(dplyr) 
library(qdap)
library(tidyr) 
library(quanteda) 
library(tidytext) 
library(reshape)
library(hunspell) 
library(udpipe) 
library(lexicon)
```

# Load data
```{r}
sample <- read_csv(file="Dataset2.csv")
```

###### Only English Reviews ######
filter so that only English reviews are included 
-> exclude 189 reviews
```{r}
sample$language <- textcat(sample$text)
sample$language <- as.factor(sample$language)
levels(sample$language)
w = table(sample$language)
w #keep english & scots
df <- sample %>%
  filter(language == c('english') | language == c('scots'))
rm(eng, sample, w, scots)
```

###### First round of cleaning ######
- Remove "\", space before a comma and extra spaces before endmarks
- Remove "<U+00E9>" which is replacement for special characters (e.g. é)
```{r}
text_df <- data_frame(df$review_id, df$text)
names(text_df)[names(text_df) == 'df$review_id'] <- 'review_id'
names(text_df)[names(text_df) == 'df$text'] <- 'text'

text_df$text <- base::gsub("<U+00E9>", "", text_df$text, fixed = TRUE)

text_df$text <- scrubber(text_df$text, 
                             num2word = FALSE, 
                             rm.quote = TRUE, 
                             fix.comma = TRUE,
                             fix.space = TRUE)
df$text[999]
text_df$text[999]
df$text[768]
text_df$text[768]
```

###### Sum of Compliments ######
- variable: total compliments a user has received
```{r}
df$reviewer_compliments <- df$compliment_cool + df$compliment_cute + df$compliment_funny + df$compliment_hot + df$compliment_list + df$compliment_more + df$compliment_note + df$compliment_plain + df$compliment_photos + df$compliment_profile + df$compliment_writer
```

# Keep only the variables that are needed
```{r}
df <- df %>% select(review_id, user_id, DV, text, stars.y, date2, review_count.y, elite, fans, yelping_since, reviewer_compliments)
```

###### Present IVs ######
- Review Star Rating: review_stars
- Count of reviews user has written: reviewer_review_count
- Count of compliments the user has received: reviewer_compliments
- Count of fans the user has: reviewer_fans
- Since when has the user been a Yelp member: reviewer_join_date
- Date the review was written: review_date
```{r}
names(df)[names(df) == 'stars.y'] <- 'review_stars'
names(df)[names(df) == 'review_count.y'] <- 'reviewer_review_count'
names(df)[names(df) == 'fans'] <- 'reviewer_fans'
names(df)[names(df) == 'yelping_since'] <- 'reviewer_join_date'
names(df)[names(df) == 'date2'] <- 'review_date'
```

###### Feature Engineering ######
Includes:
- Count of years user was Elite: reviewer_elite
- Length of review in characters with spaces: review_length_char_w
- Length of review in characters without spaces: review_length_char_wo
- Length of review in words: review_length_words
- Length of review in sentences: review_length_sentences
- Count of expressive punctuation: review_ex_punctuation
- Count of upper case letters: review_upper_letters

# Reviewer Elite
```{r}
#make sure package plyr is not loaded
df$elite <- as.character(df$elite)
elite_data <- df %>% select(user_id, elite)

elite2 <- elite_data[!duplicated(elite_data$user_id),]
head(elite2)

elite2$elite <- as.character(elite2$elite)

e <- elite2 %>%
  transform(elite = strsplit(elite, ",")) %>% unnest()
e
e2 <- e %>% group_by(user_id) %>% count()
View(e2)

e2$n[e2$n == 1] <- 0
e2

df <- merge(df, e2, by = "user_id")
rm(e)
rm(e2)
rm(elite_data)

df$elite <- NULL
df$user_id <- NULL

names(df)[names(df) == 'n'] <- 'reviewer_elite'
```

# Review length in characters
```{r}
#withouot spaces
text_df$review_length_char_wo <- character_count(text_df$text)

#with spaces
#text_df$review_length_char_w <- character_count(text_df$text,count.space = TRUE) 
#not including this one
text_df$review_length_char_w <- nchar(text_df$text) #more reliable
```

# Review length in words
```{r}
text_df$review_length_words <- sapply(gregexpr("\\S+", text_df$text), length)
text_df$review_length_words[5]
text_df$text[5]
```

# Review length in sentences
```{r}
text_df$review_length_sentences <- nsentence(text_df$text)
text_df$review_length_sentences[5]
```

# Count of expressive punctuation
- Expressive punctuation = "?" or "!"
```{r}
text_df$review_ex_punctuation <- sapply(gregexpr("[!?]", text_df$text), length)
```

# Count of upper case letters
```{r}
text_df$review_upper_letters <- sapply(regmatches(text_df$text, gregexpr("[A-Z]", text_df$text, perl=TRUE)), length)
text_df$text[4]
text_df$review_upper_letters[4]
```

# Add variables to df
```{r}
df <- merge(df, text_df, by = "review_id")
df$text.y <- NULL
```

###### BOW ######
using approach from: https://www.tidytextmining.com/index.html

```{r}
text_df$review_length_char_wo <- NULL
text_df$review_length_char_w <- NULL
text_df$review_length_words <- NULL
text_df$review_length_sentences <- NULL
text_df$review_ex_punctuation <- NULL
text_df$review_upper_letters <- NULL
head(text_df)
```

transform to a tidy data structure
-> strips punctuation
-> converts to lowercase
```{r}
tidy_text <- text_df %>%
  unnest_tokens(word, text)
head(tidy_text)
```

# Stopwords
- Number of stopwords in the review: review_stopwords
```{r}
data(stop_words)

#includes 3 dictionaries: onix, SMART and snowball
stopwords_df <- tidy_text %>%
  group_by(review_id) %>%
  inner_join(stop_words)
```

keep onix lexicon only
http://www.lextek.com/manuals/onix/stopwords1.html
```{r}
stopwords_df <- stopwords_df %>%
  filter(lexicon == "onix")
```

count of stopwords per review
```{r}
stopwords_count <- stopwords_df %>%
  group_by(review_id) %>%
  count()
```

attach to df
```{r}
df <- merge(df, stopwords_count, by = "review_id")
names(df)[names(df) == 'n'] <- 'review_stopwords'
```

clear some space
```{r}
rm(stopwords_count, stopwords_df)
```

text without stopwords
```{r}
tidy_text_wo_stop <- tidy_text %>%
  anti_join(stop_words)
```

# Unique words
- Proportion of unique words in the review, excluding stopwords: review_prop_unique
```{r}
#using data without stopwords
review_words <- tidy_text_wo_stop %>%
  count(review_id, word) %>%
  ungroup()

total_words <- review_words %>%
  group_by(review_id) %>%
  summarise(total = sum(n))
```

combine total_words with review_words
```{r}
review_words <- left_join(review_words, total_words)
```

count the number of unique words in a review
```{r}
count.unique <- review_words %>%
  filter(n == 1) %>%
  group_by(review_id) %>%
  count(n)
```

attach to total words -> divide unique by total to get proportion of unique words
```{r}
merge <- inner_join(review_words, count.unique)
merge$word <- NULL
merge$n <- NULL
merge %>%
  distinct(review_id, keep.all = TRUE)

merge <- merge %>% filter (! duplicated(review_id))
merge2 <- merge %>%
  mutate(unique.ratio = nn / total)
View(merge2)
```

attach unique word ratio to df
```{r}
merge2$total <- NULL
merge2$nn <- NULL
df <- merge(df, merge2, by = "review_id")
names(df)[names(df) == 'unique.ratio'] <- 'review_prop_unique'
names(df)[names(df) == 'text.x'] <- 'text'
colnames(df)
```

clear space
```{r}
rm(count.unique, merge, merge2, review_words, stop_words, total_words)
```

###### Readability ######
Four readability tests:
- Automated Readability Index (ARI)
- Coleman Liau (CLI)
- Flesch Kincaid
- Gunning FOG (FOG)
```{r}
readability <- textstat_readability(text_df$text, measure = c("ARI", "Coleman.Liau.grade", "Flesch.Kincaid", "FOG"), remove_hyphens = TRUE)

#attach to text_df
text_df <- cbind(text_df, readability)
text_df$document <- NULL

#attach to df
df <- merge(df, text_df, by = "review_id")
df$text.y <- NULL
names(df)[names(df) == 'ARI'] <- 'review_read_ARI'
names(df)[names(df) == 'Coleman.Liau.grade'] <- 'review_read_CLG'
names(df)[names(df) == 'Flesch.Kincaid'] <- 'review_read_FK'
names(df)[names(df) == 'FOG'] <- 'review_read_FOG'

#take the average of all four scores
df <- df %>% mutate(review_readability = (review_read_ARI + review_read_CLG + review_read_FK + review_read_FOG) / 4)
```

#clean text_df
```{r}
text_df$ARI <- NULL
text_df$Coleman.Liau.grade <- NULL
text_df$Flesch.Kincaid <- NULL
text_df$FOG <- NULL
```

###### Minimal Word Frequency Matrix ######

# One / Two / More letter words
https://cran.r-project.org/web/packages/qdap/vignettes/tm_package_compatibility.pdf

construct a minimal Word Frequency Matrix
use text_df
```{r}
qdap_wfm <- with(text_df, wfm(text, review_id))
one <- as.data.frame(qdap_wfm)
```

Filter one letter words
```{r}
one.letter.words <- Filter(qdap_wfm, min = 1, max = 1, count.apostrophe = FALSE)
one.letter.words <- as.data.frame(one.letter.words)
View(one.letter.words)
df$text.x[2]
```

Filter two letter words
```{r}
two.letter.words <- Filter(qdap_wfm, min = 2, max = 2, count.apostrophe = FALSE)
two.letter.words <- as.data.frame(two.letter.words)
```

Filter more letter words
```{r}
more.letter.words <- Filter(qdap_wfm, min = 3, count.apostrophe = FALSE)
more.letter.words <- as.data.frame(more.letter.words)
```

reshape dataframes with one two and more letter words -> use reshape package
```{r}
one.letter.words <- t(one.letter.words)
one.letter.words <- as.data.frame(one.letter.words)

two.letter.words <- t(two.letter.words)
two.letter.words <- as.data.frame(two.letter.words)

more.letter.words <- t(more.letter.words)
more.letter.words <- as.data.frame(more.letter.words)
```

clear space
```{r}
rm(one)
```

sum of one / two / more letter observations
```{r}
one.letter.words$count.one <- rowSums(one.letter.words, na.rm = FALSE, dims = 1)
two.letter.words$count.two <- rowSums(two.letter.words, na.rm = FALSE, dims = 1)
more.letter.words$count.more <- rowSums(more.letter.words, na.rm = FALSE, dims = 1)
```

convert row names into variable
```{r}
one.letter.words$review_id <- rownames(one.letter.words)
two.letter.words$review_id <- rownames(two.letter.words)
more.letter.words$review_id <- rownames(more.letter.words)
```

merge with df (and delete unnecessary columns)
```{r}
one.letter.words <- one.letter.words %>%
  select(count.one, review_id)
merge1 <- merge(df, one.letter.words, by="review_id")

two.letter.words <- two.letter.words %>%
  select(count.two, review_id)
merge2 <- merge(merge1, two.letter.words, by="review_id")

more.letter.words <- more.letter.words %>%
  select(count.more, review_id)
merge3 <- merge(merge2, more.letter.words, by="review_id")
```

clear some space
```{r}
rm(df)
rm(merge1)
rm(merge2)
rm(more.letter.words)
rm(one.letter.words)
rm(two.letter.words)
df <- merge3
rm(merge3)
```

rename new variables
```{r}
names(df)[names(df) == 'count.one'] <- 'review_one_letter_words'
names(df)[names(df) == 'count.two'] <- 'review_two_letter_words'
names(df)[names(df) == 'count.more'] <- 'review_more_letter_words'
```

# Save
```{r}
write.csv(df, file="save2.csv")
```

###### Spelling Mistakes ######
- Number of spelling mistakes in a review: review_prop_spelling_mistakes

using hunspell dictionary
```{r}
hunspell <- hunspell(text_df$text)
#hunspell
text_df$spelling.mistakes <- lengths(hunspell)
spelling <- text_df
spelling$text <- NULL
spelling$value <- NULL
```

merge with df
```{r}
df <- merge(df, spelling, by = "review_id")
```

create new variable with the ratio
```{r}
df$review_prop_spelling_mistakes <- df$spelling.mistakes / df$review_length_words
df$spelling.mistakes <- NULL
```

##### POS Tagging #####
using package udpipe
write function to apply english
```{r}
if (file.exists("english-ud-2.0-170801.udpipe")) 
  ud_model <- udpipe_load_model(file = "english-ud-2.0-170801.udpipe") else {
    ud_model <- udpipe_download_model(language = "english")
    ud_model <- udpipe_load_model(ud_model$file_model)
}
```

run function on text_df
```{r}
x <- udpipe_annotate(ud_model, text_df$text, doc_id = text_df$review_id)
x <- as.data.frame(x)

table <- table(x$doc_id, x$upos)
table
```

POS tags respectively
```{r}
table(x$upos)
x$Noun <- ifelse(x$upos == "NOUN", 1, 0)
x$Verb <- ifelse(x$upos == "VERB", 1, 0)
x$Adjective <- ifelse(x$upos == "ADJ", 1, 0)  
x$Punctuation <- ifelse(x$upos == "PUNCT", 1, 0)
x$Pronoun <- ifelse(x$upos == "PRON", 1, 0)
x$Adverb <- ifelse(x$upos == "ADV", 1, 0)
x$Numeric <- ifelse(x$upos == "NUM", 1, 0)
x$X <- ifelse(x$upos == "X", 1, 0)
x$Adpositions <- ifelse(x$upos == "ADP", 1, 0)
x$Conjunctions <- ifelse(x$upos == "CCONJ", 1, 0)
x$Determiners <- ifelse(x$upos == "DET", 1, 0)
x$Particles <- ifelse(x$upos == "PART", 1, 0)
x$Symbols <- ifelse(x$upos == "SYM", 1, 0)
```

count of POS tags respectively
```{r}
library(plyr)
x.noun = ddply(x, .(doc_id), summarise, Noun=sum(Noun), .drop=FALSE)
x.verb = ddply(x, .(doc_id), summarise, Verb=sum(Verb), .drop=FALSE)
x.adjective = ddply(x, .(doc_id), summarise, Adjective=sum(Adjective), .drop=FALSE)
x.punctuation = ddply(x, .(doc_id), summarise, Punctuation=sum(Punctuation), .drop=FALSE)
x.pronoun = ddply(x, .(doc_id), summarise, Pronoun=sum(Pronoun), .drop=FALSE)
x.adverb = ddply(x, .(doc_id), summarise, Adverb=sum(Adverb), .drop=FALSE)
x.numeric = ddply(x, .(doc_id), summarise, Numeric=sum(Numeric), .drop=FALSE)
x.X = ddply(x, .(doc_id), summarise, X=sum(X), .drop=FALSE)
x.adpositions = ddply(x, .(doc_id), summarise, Adpositions=sum(Adpositions), .drop=FALSE)
x.conjunctions = ddply(x, .(doc_id), summarise, Conjunctions=sum(Conjunctions), .drop=FALSE)
x.determiners = ddply(x, .(doc_id), summarise, Determiners=sum(Determiners), .drop=FALSE)
x.particles = ddply(x, .(doc_id), summarise, Particles=sum(Particles), .drop=FALSE)
x.symbols = ddply(x, .(doc_id), summarise, Symbols=sum(Symbols), .drop=FALSE)
```

merge into a single dataframe
```{r}
names(x.noun)[names(x.noun) == 'doc_id'] <- 'review_id'
df <- merge(x.noun, df, by = "review_id")

names(x.verb)[names(x.verb) == 'doc_id'] <- 'review_id'
df <- merge(x.verb, df, by = "review_id")

names(x.adjective)[names(x.adjective) == 'doc_id'] <- 'review_id'
df <- merge(x.adjective, df, by = "review_id")

names(x.punctuation)[names(x.punctuation) == 'doc_id'] <- 'review_id'
df <- merge(x.punctuation, df, by = "review_id")

names(x.pronoun)[names(x.pronoun) == 'doc_id'] <- 'review_id'
df <- merge(x.pronoun, df, by = "review_id")

names(x.adverb)[names(x.adverb) == 'doc_id'] <- 'review_id'
df <- merge(x.adverb, df, by = "review_id")

names(x.numeric)[names(x.numeric) == 'doc_id'] <- 'review_id'
df <- merge(x.numeric, df, by = "review_id")

names(x.X)[names(x.X) == 'doc_id'] <- 'review_id'
df <- merge(x.X, df, by = "review_id")

names(x.adpositions)[names(x.adpositions) == 'doc_id'] <- 'review_id'
df <- merge(x.adpositions, df, by = "review_id")

names(x.conjunctions)[names(x.conjunctions) == 'doc_id'] <- 'review_id'
df <- merge(x.conjunctions, df, by = "review_id")

names(x.determiners)[names(x.determiners) == 'doc_id'] <- 'review_id'
df <- merge(x.determiners, df, by = "review_id")

names(x.particles)[names(x.particles) == 'doc_id'] <- 'review_id'
df <- merge(x.particles, df, by = "review_id")

names(x.symbols)[names(x.symbols) == 'doc_id'] <- 'review_id'
df <- merge(x.symbols, df, by = "review_id")
```

rename variables
```{r}
names(df)[names(df) == 'Noun'] <- 'review_nouns'
names(df)[names(df) == 'Verb'] <- 'review_verbs'
names(df)[names(df) == 'Adjective'] <- 'review_adjectives'
names(df)[names(df) == 'Punctuation'] <- 'review_punctuation'
names(df)[names(df) == 'Pronoun'] <- 'review_pronouns'
names(df)[names(df) == 'Adverb'] <- 'review_adverbs'
names(df)[names(df) == 'Numeric'] <- 'review_numbers'
names(df)[names(df) == 'X'] <- 'review_X'
names(df)[names(df) == 'Adpositions'] <- 'review_adpositions'
names(df)[names(df) == 'Conjunctions'] <- 'review_conjunctions'
names(df)[names(df) == 'Determiners'] <- 'review_determiners'
names(df)[names(df) == 'Particles'] <- 'review_particles'
names(df)[names(df) == 'Symbols'] <- 'review_symbols'
```

#Clear space
```{r}
rm(elite2, spelling, x.adjective, x.noun, x.numeric, x.verb, x.punctuation, x.adpositions, x.conjunctions, x.determiners, x.particles, x.X)
rm(x.adverb, x.pronoun)
```

###### Semantic Analysis ######
- review polarity: review_polarity

Use spread() so that negative and positive sentiment are in separate columns, then calculate a net sentiment (pos - neg) -> Using "bing" lexicon
```{r}
#unload package plyr
library(tidyr)
library(dplyr)
library(tidytext)
review_polarity <- tidy_text %>%
  left_join(get_sentiments("bing")) %>%
  count(review_id, sentiment) %>%
  spread(sentiment, n, fill = 0, drop = FALSE) %>%
  mutate(sentiment = positive - negative)
```

attach sentiment to df
```{r}
df <- merge(df, review_polarity, by = "review_id")
df$negative <- NULL
df$positive <- NULL
names(df)[names(df) == 'sentiment'] <- 'review_polarity'
```

#### CHECK ####
```{r}
colnames(df)
```

# Filter only the features used in analysis
```{r}
df <- df %>% select(DV, review_symbols, review_particles, review_determiners, review_conjunctions, review_adpositions, review_numbers, review_adverbs, review_pronouns, review_punctuation, review_adjectives, review_verbs, review_nouns, review_stars, review_date, review_length_char_wo, review_length_char_w, review_length_words, review_length_sentences, review_ex_punctuation, review_upper_letters, review_stopwords, review_prop_unique, review_readability, review_one_letter_words, review_two_letter_words, review_more_letter_words, review_prop_spelling_mistakes, review_polarity, reviewer_review_count, reviewer_fans, reviewer_join_date, reviewer_compliments, reviewer_elite)

colnames(df)
```

#### PROPORTIONS #####
-> standardize

# With total number of words
```{r}
#POS
df$review_prop_nouns <- df$review_nouns / df$review_length_words
df$review_nouns <- NULL

df$review_prop_verbs <- df$review_verbs / df$review_length_words
df$review_verbs <- NULL

df$review_prop_adverbs <- df$review_adverbs / df$review_length_words
df$review_adverbs <- NULL

df$review_prop_pronouns <- df$review_pronouns / df$review_length_words
df$review_pronouns <- NULL

df$review_prop_adjectives <- df$review_adjectives / df$review_length_words
df$review_adjectives <- NULL

df$review_prop_particles <- df$review_particles / df$review_length_words
df$review_particles <- NULL

df$review_prop_determiners <- df$review_determiners / df$review_length_words
df$review_determiners <- NULL

df$review_prop_conjunctions <- df$review_conjunctions / df$review_length_words
df$review_conjunctions <- NULL

df$review_prop_adpositions <- df$review_adpositions / df$review_length_words
df$review_adpositions <- NULL

df$review_prop_numbers <- df$review_numbers / df$review_length_words
df$review_numbers <- NULL

#Stopwords
df$review_prop_stopwords <- df$review_stopwords / df$review_length_words
df$review_stopwords <- NULL

#One / two / more letter words
df$review_prop_one_letter_words <- df$review_one_letter_words / df$review_length_words
df$review_one_letter_words <- NULL

df$review_prop_two_letter_words <- df$review_two_letter_words / df$review_length_words
df$review_two_letter_words <- NULL

df$review_prop_more_letter_words <- df$review_more_letter_words / df$review_length_words
df$review_more_letter_words <- NULL
```

# With total number of characters excluding white spaces
```{r}
df$review_prop_symbols <- df$review_symbols / df$review_length_char_wo
df$review_symbols <- NULL

df$review_prop_punctuation <- df$review_punctuation / df$review_length_char_wo
df$review_punctuation <- NULL

df$review_prop_ex_punctuation <- df$review_ex_punctuation / df$review_length_char_wo
df$review_ex_punctuation <- NULL

df$review_prop_uppercase <- df$review_upper_letters / df$review_length_char_wo
df$review_upper_letters <- NULL
```

# Check
```{r}
colnames(df)
summary(df)
```

# SAVE
```{r}
write.csv(df, file="DatasetFinal.csv")
```



